---
title: "Demographic response of a dioecious range-limited species to climate change"
author: "Tom Miller,Jacob Moutouama and Aldo Compagnoni"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
 rmdformats::readthedown:
    code_folding: show
    self_contained: true
    number_sections: true
    thumbnails: true
    lightbox: true
    gallery: true
    keep_md: true
    highlight: tango
    df_print: kable 
    toc_depth: 3
    fig_width: 8
    fig_height: 8
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, 
                      echo = TRUE,
                      warning=FALSE, 
                      message=FALSE)
knitr::opts_knit$set(global.par = TRUE,knitr.table.format = "latex")
```

```{r eval=TRUE,echo=T}
rm(list = ls())

```

# Required libraries
```{r}
# load packages
library(rstan)
# set rstan options
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
set.seed(13)
# Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(tidyverse.quiet = TRUE)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(bayesplot)
# install.packages("countreg",repos = "http://R-Forge.R-project.org")
library(countreg)
library(rmutil)
library(actuar)
library(SPEI)
library(LaplacesDemon)
```

# Demography and germination data  
```{r}
poar_allsites <- read.csv("https://www.dropbox.com/s/xk4225mn8btqhbm/demography_allsites.csv?dl=1", stringsAsFactors = F)#common garden data
viabVr <- read.csv("https://www.dropbox.com/s/jfkgoxgv8o1fgqx/viability.csv?dl=1") #seed viability and germination
```

# Add current climatic conditions to match with the demograhic data
```{r}
poar_allsites$census.year<-poar_allsites$year-1 #Add census year to match with climate data 
summary(poar_allsites)

poar_ppt <- read.csv("/Users/jm200/Library/CloudStorage/Box-Box/Miller Lab/Moutouama Rice University/Forecasting/data/Poa_pr.csv", stringsAsFactors = F) # monthly  precipitation 
poar_temp <- read.csv("/Users/jm200/Library/CloudStorage/Box-Box/Miller Lab/Moutouama Rice University/Forecasting/data/Poa_tas.csv", stringsAsFactors = F) # monthly temperature)
poar_ppt$census.year<-ifelse(poar_ppt$month<5,poar_ppt$year-1,poar_ppt$year) # data were collected in May. Thus, the precipitation for the census year is both the precipitation from May to December of the previous year and precipitation of the current year.  
poar_temp$census.year<-ifelse(poar_temp$month<5,poar_temp$year-1,poar_temp$year)

poar_ppt %>% 
  dplyr::select(Longitude, Latitude, census.year, ppt,site,long.scaled,long.center) %>% 
  filter(census.year %in% (2014:2016)) %>%
  group_by(site,census.year) %>%
	summarise(mean_ppt=sum(ppt),sd_ppt=sd(ppt),cv_ppt=sd_ppt/mean_ppt,Longitude=unique(Longitude),Latitude=unique(Latitude))->poar_pptann # annual precipitation and seasonality of precipitation seasonality of temperature

poar_temp %>% 
  dplyr::select(Longitude, Latitude, census.year, temp,site,long.scaled,long.center) %>% 
  filter(census.year %in% (2014:2016)) %>%
  group_by(site,census.year) %>%
	summarise(mean_temp = mean(temp),sd_temp=sd(temp),cv_temp=sd_temp/mean_temp,Longitude=unique(Longitude),Latitude=unique(Latitude))->poar_tempann # mean annual temperature and 
poar_clim<- merge(x = poar_pptann, y = poar_tempann) # merge temperature and precipitation data

poar_allsites.clim <- left_join(x = poar_allsites ,y =poar_clim,by=c("site","census.year","Longitude","Latitude")) # merge the demographic data with climatic data for each site
poar_allsites.clim$siteyear<-interaction(poar_allsites.clim$site,poar_allsites.clim$census.year) # create a new variable taking into account the year and the climate variable
poar_allsites.clim$siteyear<-as.character(poar_allsites.clim$siteyear)

```

# Check for correlation between predictors ( precipitation and seasonality of precipitation,  temperature and seasonality of temperature )

```{r,eval=FALSE}
figppt <- ggpubr::ggscatter(poar_allsites.clim, x = "mean_ppt", y = "cv_ppt",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
figppt + ggpubr::stat_cor(method = "pearson", label.x = 1000, label.y = 0.09)
```

```{r,eval=FALSE}
fitemp <- ggpubr::ggscatter(poar_allsites.clim, x = "mean_temp", y = "cv_temp",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
fitemp + ggpubr::stat_cor(method = "pearson", label.x = 14, label.y = 0.2)
```
# Data for each vital rate
```{r}
# Survival
poar_allsites.clim %>% 
  subset( tillerN_t0 > 0 )%>%
  dplyr::select( year, Code, site, Block, Sex, 
          Longitude, Latitude, 
          tillerN_t0, surv_t1,mean_ppt,cv_ppt,mean_temp,siteyear)%>% 
  na.omit %>% 
  mutate( site         = site %>% as.factor %>% as.numeric,
          Block = Block %>% as.factor %>% as.numeric,
          Sex          = Sex %>% as.factor %>% as.numeric,
          siteyear = siteyear %>% as.factor %>% as.numeric,
          source = Code %>% as.factor %>% as.numeric ) %>%
  mutate( log_size_t0   = log(tillerN_t0),
          log_ppt_t1_z = log(mean_ppt) %>% scale %>% .[,1],
          log_cvppt_t1_z = log(cv_ppt) %>% scale %>% .[,1],
          log_temp_t1_z = log(mean_temp) %>% scale %>% .[,1])->poar_allsites.surv
# growth
poar_allsites.clim %>% 
  subset( tillerN_t0 > 0 & tillerN_t1 > 0)%>%
  dplyr::select( year, Code, site, Block, Sex, 
          Longitude, Latitude, 
          tillerN_t0, tillerN_t1,surv_t1,mean_ppt,cv_ppt,mean_temp,siteyear)%>% 
  na.omit %>% 
  mutate( site         = site %>% as.factor %>% as.numeric,
          Block = Block %>% as.factor %>% as.numeric,
          Sex          = Sex %>% as.factor %>% as.numeric,
          siteyear = siteyear %>% as.factor %>% as.numeric,
          source = Code %>% as.factor %>% as.numeric ) %>%
  mutate( log_size_t0   = log(tillerN_t0),
          log_ppt_t1_z = log(mean_ppt) %>% scale %>% .[,1],
          log_cvppt_t1_z = log(cv_ppt) %>% scale %>% .[,1],
          log_temp_t1_z = log(mean_temp) %>% scale %>% .[,1])->poar_allsites.grow
# flowering
poar_allsites.clim %>% 
  subset( tillerN_t1 > 0 )%>%
  dplyr::select( year, Code, site, Block, Sex, 
          Longitude, Latitude, 
          tillerN_t1, flowerN_t1,flow_t1,mean_ppt,cv_ppt,mean_temp,siteyear)%>% 
  na.omit %>% 
  mutate( site         = site %>% as.factor %>% as.numeric,
          Block = Block %>% as.factor %>% as.numeric,
          Sex          = Sex %>% as.factor %>% as.numeric,
          siteyear = siteyear %>% as.factor %>% as.numeric,
          source = Code %>% as.factor %>% as.numeric ) %>%
  mutate( log_size_t1   = log(tillerN_t1),
          log_ppt_t1_z = log(mean_ppt) %>% scale %>% .[,1],
          log_cvppt_t1_z = log(cv_ppt) %>% scale %>% .[,1],
          log_temp_t1_z = log(mean_temp) %>% scale %>% .[,1])->poar_allsites.flow
# Panicles
poar_allsites.clim %>% 
  subset( flowerN_t1 > 0 & tillerN_t1 > 0 )%>%
  dplyr::select( year, Code, site, Block, Sex, 
          Longitude, Latitude, 
          tillerN_t1, flowerN_t1,mean_ppt,cv_ppt,mean_temp,siteyear)%>% 
  na.omit %>% 
  mutate( site         = site %>% as.factor %>% as.numeric,
          Block = Block %>% as.factor %>% as.numeric,
          Sex          = Sex %>% as.factor %>% as.numeric,
          siteyear = siteyear %>% as.factor %>% as.numeric,
          source = Code %>% as.factor %>% as.numeric,
          panic_t1 = flowerN_t1) %>%
  mutate( log_size_t1   = log(tillerN_t1),
          log_ppt_t1_z = log(mean_ppt) %>% scale %>% .[,1],
          log_cvppt_t1_z = log(cv_ppt) %>% scale %>% .[,1],
          log_temp_t1_z = log(mean_temp) %>% scale %>% .[,1])->poar_allsites.panic

# seed viability
viabVr %>% 
  dplyr::select( plot, totS, yesMaybe, sr_f ) %>% 
  rename( SR        = sr_f,
          y_viab = yesMaybe,
          tot_seeds_viab = totS) %>% 
  dplyr::select(y_viab, tot_seeds_viab, SR ) %>% 
  na.omit->viab

# seed germination
viabVr %>% 
  dplyr::select( plot, germTot, germFail, sr_f ) %>% 
  rename( SR        = sr_f,
          y_germ    = germTot ) %>% 
  mutate(tot_seeds_germ = y_germ + germFail ) %>% 
  dplyr::select(y_germ, tot_seeds_germ, SR ) %>% 
  na.omit->germ

# seeds per panicle
viabVr %>% 
  dplyr::select(SeedN)  %>% 
  na.omit->seeds

```

# Bundle data for model for current climatic conditions
```{r}

data_allsites <- list( n_sites    = poar_allsites.surv$site %>% n_distinct,
                  n_sources  = poar_allsites.surv$source %>% n_distinct(),
                  
                  # survival data
                  n_blocks_s = poar_allsites.surv$Block %>% n_distinct,
                           n_siteyear_s = poar_allsites.surv$siteyear %>% n_distinct,
                  site_s   = poar_allsites.surv$site,
                  source_s =  poar_allsites.surv$source,
                  block_s  = poar_allsites.surv$Block,
                  siteyear_s  = poar_allsites.surv$siteyear,
                  ppt_s=poar_allsites.surv$log_ppt_t1_z,
                  cvppt_s=poar_allsites.surv$log_cvppt_t1_z,
                  temp_s=poar_allsites.surv$log_temp_t1_z,
                  site_block_s = data.frame( site_i  = poar_allsites.surv$site,
                                             block_i = poar_allsites.surv$Block ) %>%
                    unique %>% .$site_i,
                  male_s   = poar_allsites.surv$Sex-1,
                  size_s   = poar_allsites.surv$log_size_t0,
                  y_s      = poar_allsites.surv$surv_t1,
                  n_s      = nrow(poar_allsites.surv),
                  
                  # growth data
                  n_blocks_g = poar_allsites.grow$Block %>% n_distinct,
                           n_siteyear_g = poar_allsites.grow$siteyear %>% n_distinct,
                  site_g   = poar_allsites.grow$site,
                  source_g =  poar_allsites.grow$source,
                  block_g  = poar_allsites.grow$Block,
                  siteyear_g  = poar_allsites.grow$siteyear,
                  ppt_g=poar_allsites.grow$log_ppt_t1_z,
                  cvppt_g=poar_allsites.grow$log_cvppt_t1_z,
                  temp_g=poar_allsites.grow$log_temp_t1_z,
                  site_block_g = data.frame( site_i  = poar_allsites.grow$site,
                                             block_i = poar_allsites.grow$Block ) %>%
                    unique %>% .$site_i,
                  male_g   = poar_allsites.grow$Sex-1,
                  size_g   = poar_allsites.grow$log_size_t0,
                  y_g      = poar_allsites.grow$tillerN_t1,
                  n_g      = nrow(poar_allsites.grow),

                  # flowering data
                   n_blocks_f = poar_allsites.flow$Block %>% n_distinct,
                           n_siteyear_f = poar_allsites.flow$siteyear %>% n_distinct,
                  site_f   = poar_allsites.flow$site,
                  source_f =  poar_allsites.flow$source,
                  block_f  = poar_allsites.flow$Block,
                  siteyear_f  = poar_allsites.flow$siteyear,
                  ppt_f=poar_allsites.flow$log_ppt_t1_z,
                  cvppt_f=poar_allsites.flow$log_cvppt_t1_z,
                  temp_f=poar_allsites.flow$log_temp_t1_z,
                  site_block_f = data.frame( site_i  = poar_allsites.flow$site,
                                             block_i = poar_allsites.flow$Block ) %>%
                    unique %>% .$site_i,
                  male_f   = poar_allsites.flow$Sex-1,
                  size_f   = poar_allsites.flow$log_size_t1,
                  y_f      = poar_allsites.flow$flow_t1,
                  n_f      = nrow(poar_allsites.flow),

                  
                  # panicle data
                   n_blocks_p = poar_allsites.panic$Block %>% n_distinct,
                           n_siteyear_p = poar_allsites.panic$siteyear %>% n_distinct,
                  site_p   = poar_allsites.panic$site,
                  source_p =  poar_allsites.panic$source,
                  block_p  = poar_allsites.panic$Block,
                  siteyear_p  = poar_allsites.panic$siteyear,
                  ppt_p=poar_allsites.panic$log_ppt_t1_z,
                  cvppt_p=poar_allsites.panic$log_cvppt_t1_z,
                  temp_p=poar_allsites.panic$log_temp_t1_z,
                  site_block_p = data.frame( site_i  = poar_allsites.panic$site,
                                             block_i = poar_allsites.panic$Block ) %>%
                    unique %>% .$site_i,
                  male_p   = poar_allsites.panic$Sex-1,
                  size_p   = poar_allsites.panic$log_size_t1,
                  y_p      = poar_allsites.panic$panic_t1,
                  n_p      = nrow(poar_allsites.panic), 

                  
                  # viability
                  n_v       = nrow(viab),
                  y_v       = viab$y_viab,
                  tot_seeds_v = viab$tot_seeds_viab,
                  SR_v        = viab$SR,
                  
                  # germination
                  n_m       = nrow(germ),
                  y_m       = germ$y_germ,
                  tot_seeds_m = germ$tot_seeds_germ,
                  SR_m        = germ$SR,
                  
                  # seeds per panicle
                  n_d = nrow(seeds),
                  y_d = seeds$SeedN)
```

# MCMC parameters
```{r}
sim_pars <- list(
  warmup = 1000,
  iter = 4000,
  thin = 3,
  chains = 3
)

```

```{r flowering, cache=TRUE}
fit_allsites_full <- stan(
 file = "/Users/jm200/Library/CloudStorage/Box-Box/Miller Lab/Moutouama Rice University/Forecasting/stan/Code/poar_viability.stan",
 data = data_allsites,
 warmup = sim_pars$warmup,
 iter = sim_pars$iter,
 thin = sim_pars$thin,
 chains = sim_pars$chains,
 seed=13)
# saveRDS(fit_allsites_full, 'C:/YOURDIRECTORY')

```

# Diagnostics
## Posterior Distributions
```{r}
traceplot(fit_allsites_full, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bpptsex_p","btempsex_p","bcvpptsex_p","bppttempsex_p"), inc_warmup = TRUE, nrow = 2) # The traceplot method is used to plot the time series of the posterior draws. If we include the warmup draws by setting inc_warmup=TRUE, the background color of the warmup area is different from the post-warmup phase

```

```{r}
traceplot(fit_allsites_full, pars = c("v0","a_v","m","phi_m"), inc_warmup = TRUE, nrow = 2) # The traceplot method is used to plot the time series of the posterior draws. If we include the warmup draws by setting inc_warmup=TRUE, the background color of the warmup area is different from the post-warmup phase

```



```{r}
pairs(fit_allsites_paniclecv, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p"), las = 2)# We can also make a graphical representation of (much of the) the same information using pairs. The “pairs”" plot can be used to get a sense of whether any sampling difficulties are occurring in the tails or near the mode. Here btempptt has an issue. The marginal distribution of each selected parameter is included as a histogram along the diagonal.Ideally, the below-diagonal intersection and the above-diagonal intersection of the same two variables should have distributions that are mirror images of each other. Any yellow points would indicate transitions where the maximum treedepth__ was hit, and red points indicate a divergent transition.

```


```{r}
stan_ac(fit_allsites_paniclecv, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p"), lags = 15)# We can also evaluate the autocorrelation between samples for each parameter at different lags. Given the nature of Markov chains, there has to be some autocorrelation at the first lag, but we want to see a sharp drop off as the lags increase. Since we are performing this operation on multiple chains, these results are the average autocorrelation values from all four chains.

```

```{r}
coda::gelman.plot(As.mcmc.list(fit_allsites_paniclecv, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p")), autoburnin = F) # What if we want a plot of the potential scale reduction factor R̂ at each iteration of the sampler? Unfortunately, rstan doesn’t include a builtin function to produce this plot. Fortunately, it does provide an incredibly easy way to convert a stanfit object to a coda object using the As.mcmc.list() function.
```

```{r}
stan_dens(fit_allsites_paniclecv, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p")) + geom_vline(xintercept = 0, linetype = 2)
```

```{r}
# extract regression parameters
regpar <- summary(fit_allsites_paniclecv, pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p"))$summary
knitr::kable(regpar)
```


```{r}
plot(fit_allsites_paniclecv,pars = c("b0_p", "bsize_p","bsex_p","bppt_p","btemp_p","bcvppt_p","bcvtemp_p","bpptsex_p","btempsex_p","bcvpptsex_p","bcvtempsex_p","bppttempsex_p","bcvpptcvtempsex_p"))+ geom_vline(xintercept = 0, linetype = 2)
```



# Posterior predictive check


```{r,eval=FALSE}
#Now we’ve run the models, let’s do some posterior predictive checks. The idea of posterior predictive checks is to compare our observed data to replicated data from the model. If our model is a good fit, we should be able to use it to generate a dataset that resembles the observed data.
#pull out parameter posterior distributions
predP <- rstan::extract(fit_allsites_paniclecv, pars = c("predP"))$predP
#draw 500 random samples from the joint posterior
n_post_draws <- 100
post_draws <-sample.int(dim(predP)[1],n_post_draws)
#set up simulation output
y_p_sim <- matrix(NA,n_post_draws,length(poar_allsites.panic$y_p))
#loop over the posterior and generate new observations
for(i in 1:n_post_draws){
  print(i)
  ## sample flowering data (bernoulli)
  y_p_sim[i,] <- rbinom(n=length(poar_allsites.panic$y_p), size=1, prob = invlogit(predP[i,]))
}

```


```{r,eval=FALSE}
#plot ppc overlay
(ppc_dens_overlay(poar_allsites.panic$y_p, y_p_sim)+
  xlab("  P status")+ylab("Density")+
  ggtitle(("P"))+theme(legend.position = "none")->ppc_pa)
```

